The Ex-Ante View of Recommender System Design

### 2019-09-15 paper/filter_bubbles.tex
fixups to section 4

All approaches to serendipity to our knowledge, however, do not take into account the \textit{beliefs} that users have over the valuations for the items. For instance, \cite{vargas2011rank, kaminskas2014measuring} propose a unexpectedness metric that looks at the dissimilarity of the recommended items to what the recommender already knows the user likes either via a content-based approach or a collaborative-based approach. \cite{adamopoulos2015unexpectedness} use an expected utility approach but measure unexpectedness in terms of deviations from their ``expected set of items."

We propose a definition of serendipity based on \textit{unexpectedness} and \textit{usefulness} which have have clear definitions in the context of our model of user decision-making. Users have \textit{beliefs} over the valuations for the items and thus \textit{unexpected} recommendations are given by items that consumers have low expected utility for, either due to the fact that their mean belief is low or the degree of uncertainty is high. Recall from Findings \ref{finding_local_consumption} and \ref{finding_diversity_welfare_corr} that, particularly if users are risk-averse, they will consume content in increasingly narrow portions of the product space and this may be detrimental to their realized welfare. As a result, identifying the portions of the product space where users may have significant uncertainty and utilizing recommendation to reduce this uncertainty can be welfare-enhancing for consumers. This motivates a definition of \textit{useful} recommendations as recommendations that give consumers information that would \textit{change the item they consume relative to no recommendation}.

Operationalizing this approach requires the collection of data that is not traditionally collected for recommender system design. In particular, it becomes important to understand what choices users would make \textit{without recommendation} and to collect data on user beliefs. \cite{jiang2014choice, saavedra2016choice} argue that choice-based approaches alone can help in designing better RS, though they argue for these approaches because they do a better job at providing more accurate recommendations. However, we argue that choice-based approaches are useful not as recommendations but as a baseline for what recommendations would be useful.\footnote{Furthermore, user beliefs and user choices contain information that may not have been observed by the recommender but is observed by the user. Ratings, reviews, friends, there are many sources of information affecting a person's beliefs and choices that are unobservable by recommender systems. Other consumption choices, e.g. movies seen in the cinema, change beliefs about for instance how good a director is, which affect beliefs about other movies and guide choices on a streaming platform that is unable to observe these data. However, this information can be inferred by the platform by collecting both beliefs and consumption choices.}

>>>>

All approaches to serendipity to our knowledge, however, do not take into account the \textit{beliefs} that users have over the valuations for the items. For instance, \cite{vargas2011rank, kaminskas2014measuring} propose unexpectedness metrics that look at the dissimilarity of the recommended items to what the recommender already knows the user likes either via a content-based approach or a collaborative-based approach. \cite{adamopoulos2015unexpectedness} use an expected utility approach but measure unexpectedness in terms of deviations from the ``expected consideration set" of a user and do not explicitly incorporate user beliefs.

We propose a definition of serendipity based on \textit{unexpectedness} and \textit{usefulness} which have have clear definitions in the context of our model of user decision-making. Users have \textit{beliefs} over the valuations for the items and thus \textit{unexpected} recommendations are given by items that consumers have low expected utility for, either due to the fact that their mean belief is low or the degree of uncertainty is high. Recall from Findings \ref{finding_local_consumption} and \ref{finding_diversity_welfare_corr} that, particularly if users are risk-averse, they will consume content in increasingly narrow portions of the product space and this may be detrimental to their realized welfare. As a result, identifying the portions of the product space where users may have significant uncertainty and utilizing recommendation to reduce this uncertainty can be welfare-enhancing for consumers. This motivates a definition of \textit{useful} recommendations as recommendations that give consumers information that would \textit{change the item they consume relative to no recommendation}. Combining these, serendipitious recommendations are recommendations that provide users with information about \textit{unexpected} items that end up being \textit{useful} by changing consumption choices to increase the realized utility of a user.

Operationalizing this approach requires the collection of data that is not traditionally collected for recommender systems. In particular, it becomes important to understand what choices users would make \textit{without recommendation} and to collect data on user beliefs, which can partially be inferred by observed choices. \cite{jiang2014choice, saavedra2016choice} argue that choice-based approaches alone can help in designing better RS, though they argue for these approaches because they do a better job at providing more accurate recommendations. However, we argue that choice-based approaches are useful not as recommendations but as a baseline for what recommendations would be useful.\footnote{Furthermore, user beliefs and user choices contain information that may not have been observed by the recommender but is observed by the user. Ratings, reviews, friends, there are many sources of information affecting a person's beliefs and choices that are unobservable by recommender systems. Other consumption choices, e.g. movies seen in the cinema, change beliefs about for instance how good a director is, which affect beliefs about other movies and guide choices on a streaming platform that is unable to observe these data. However, this information can be inferred by the platform by collecting both beliefs and consumption choices.}
